---
title: "Correlation between other Passenger Demographics (Age and Gender) and Survival Status on the Titanic"
---
```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary libraries
library(ggplot2)
library(dplyr)
# Read the Titanic dataset
data <- read.csv("titanic-data.csv")
```
```{r}
#| echo: false
#| message: false
#| warning: false
# Group data by gender and survival status, then count the occurrences
gender_survival_counts <- data %>%
  group_by(Sex, Survived) %>%
  summarise(Count = n()) %>%
  ungroup()

# Convert the 'Survived' column to a factor with labels for better readability
gender_survival_counts$Survived <- factor(gender_survival_counts$Survived, levels = c(0, 1), labels = c("Did not survive", "Survived"))

# Create a bar plot to visualize survival counts by gender
ggplot(gender_survival_counts, aes(x = Sex, y = Count, fill = Survived)) +
  # Use an identity stat to plot the pre-summarized counts and dodge position to place bars side by side
  geom_bar(stat = "identity", position = position_dodge()) +
  # changed from stacked bar plot to dodged plot to better display the difference in numbers by gender, instead of stacking, each category is shown side by side so it is more comprehensible
  # Add titles and labels to the plot
  labs(title = "Survival Counts by Gender on the Titanic", 
       x = "Gender", 
       y = "Count", 
       fill = "Survival Stat") +
  # Customize the fill colors for the survival status
  scale_fill_manual(values = c("#ffbfbf", "#bfcfff")) +
  # Apply a minimal theme to the plot
  theme_minimal()
```
The number of survivors was greater for females than males, and the number of deaths was much higher for males. This corresponds to well-known historical information, as women were prioritized when leaving the ship, resulting in more survivals for females and fewer for males.
```{r}
#| echo: false
#| message: false
#| warning: false
# Create a histogram to visualize the age distribution by survival status
ggplot(data, aes(x = Age, fill = as.factor(Survived))) +
  # Add histogram bars with a bin width of 5 years, setting transparency to 0.5 and overlapping the bars
  geom_histogram(binwidth = 5, alpha = 0.5, position = "identity") +
  # Add titles and labels to the plot
  labs(title = "Age Distribution by Survival Status on the Titanic", 
       x = "Age", 
       y = "Count") +
  # Customize the fill colors for the survival status
  scale_fill_manual(values = c("#ffbfbf", "#bfcfff"), 
                    labels = c("Did not survive", "Survived"), 
                    name = "Survival Status") +
  # Apply a minimal theme to the plot
  theme_minimal()
```
Survival numbers were the highest for ages 20 to 40, with significant numbers also evident in ages 0 to 10 and 40 to 50, although the ages with higher survival numbers also had higher death counts.
